---
title: "ESM_201_Assignment_1"
author: "Sandra Fogg"
date: "2/3/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, echo=FALSE}
# Load packages
library(tidyverse)
library(corrplot)
library(janitor)
library(lubridate)
library(ggimage)
library(kableExtra)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(extrafont)
library(ggrepel)
library(vcdExtra)
library(effsize)
library(leaflet)
library(car)
library(stargazer)
library(descr)
library(magrittr)
library(tinytex)
```

```{r, include=FALSE, echo=FALSE}
# Read csv data
pop_data_df <- read_csv("assig1_data.csv")
pop_data_df
pop_data_mills_df <- pop_data_df %>% 
  mutate(pop_mills = population/1000000)
pop_data_mills_df
```


##Question 1

###A. Describe how annual population data are used to calculate $\frac{dN}{Ndt}$. 
The expression $\frac{dN}{Ndt}$ represents the per capita rate change of a population (r) at time t. By taking the natural log of the annual population at time (t+1), dividing it by the initial number of individuals in the same population at time (t), and multiplying it by the expression $\frac{1}{(t+1)-t}$, the annual population data can be used to estimate per capita rate of change of the population. In this case, the time difference is 1 year, so the second term of the equation is equal to 1.  

###B. Provide the calculation used to determine $\frac{dN}{Ndt}$ for each country during the time period 1963 to 1964.

##### Brazil, $r= ln\left(\frac{81972001}{79602001}\right)$

##### India, $r= ln\left(\frac{486639001}{476632001}\right)$

##### Japan, $r= ln\left(\frac{96959001}{95929001}\right)$

##### Mexico, $r= ln\left(\frac{43052001}{41715001}\right)$

##### South Korea, $r= ln\left(\frac{27767001}{27138001}\right)$

```{r, include = FALSE, echo = FALSE}
brazil_r <- log(81972001/79602001)
brazil_r

india_r <- log(486639001/476632001)
india_r

japan_r <- log(96959001/95929001)
japan_r

mexico_r <- log(43052001/41715001)
mexico_r

southkorea_r <- log(27767001/27138001)
southkorea_r

set_colnames <- 

r_table <- pop_data_mills_df %>% 
  filter(year %in% c(1963,1964)) %>% 
  select(country, year, pop_mills) %>% 
  spread(year, pop_mills) %>% 
  set_colnames(value = c("Country", "N0", "Nt")) %>% 
  mutate(dNNdt_calc = log(Nt/N0),
         dNNdt = c(0.029338507, 0.020777870, 0.010679874, 0.031547910, 0.022913305))


r_table
```

```{r, echo= FALSE}
rate_table  <- r_table %>%
  select(Country, dNNdt_calc, dNNdt)
kable(rate_table, digits = 4, col.names = c("Country",  "Calculated dN/Ndt", "Actual dN/Ndt"),
      caption = "Calculated vs. Actual Per Capita Rate of Population Change by Country from 1963 to 1964") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "bordered")) 


```
The calculated values are exactly the same as the provided values for $\frac{dN}{Ndt}$

###C. Calculate population rate change of India from 1963 - 2004
```{r, echo = FALSE, message = FALSE, warning = FALSE}
india_plot <- read_csv("india_data.csv") %>% 
  mutate(pop_mill_i = population/1000000) %>%
  select(country, year, pop_mill_i,dNNdti, dNNdt_calci) %>% 
  ggplot(aes(x = dNNdt_calci, y = dNNdti)) +
  geom_point(color = "black") +
  geom_abline(method = "lm", se=FALSE, size = 0.5, color = "darkgreen") +
  labs(x = "Calculated Per Capita Rate of Change", y = "Actual Per Capita Rate of Change") +
  scale_x_continuous(limits = c(0.014, 0.024), breaks = seq(0.014, 0.024, by =0.0025)) + scale_y_continuous(limits = c(0.014, 0.024), breaks = seq(0.014, 0.024, by =0.0025)) +
  theme_classic() +
  ggtitle(expression(paste("Actual vs. Calculated Per Capita Rate of Change in India 1963-2004  ",  (frac(dN, Ndt)))))

india_plot
  
```


##QUESTION 2

###A. Graph $\frac{dN}{Ndt}$ vs. N~t~ for each country
```{r, echo=FALSE}
country_plots <- ggplot(pop_data_mills_df, aes(x = pop_mills, y= dNNdt)) +
  geom_line()+
  facet_wrap(~country, scales = "free") + 
  labs(y = expression(paste("Per Capita Rate of Change ", (frac(dN, Ndt)))), x = "Population Size (Nt, Millions)") +
  theme_classic() +
  ggtitle("Per Capita Growth Rate versus Population Density of Five Countries")

country_plots
```

###B. Estimate the carrying capacity for India and Brazil
```{r, include=FALSE}
#india_df_pop_mills <- read_csv("india_data.csv") %>% 
 # mutate(india_pm = population/1000000) %>% 
#  select(country, year, india_pm, dNNdti)
#indiaplots <- ggplot(india_df_pop_mills, aes(x = india_pm, y= dNNdti)) +
#  geom_line()+
#  labs(y = expression(paste("Per Capita Rate of Change ", frac(dN, Ndt))), x = "Population Size (Millions)") +
#  theme_classic() +
#  ggtitle("Per Capita Growth Rate versus Population Density of India") +
#  geom_smooth(method = "lm", se = FALSE) +
#  scale_x_continuous(expand = c(0,0), limits = c(500, 3000 ), breaks = seq(500, 3000, 500)) +
 # scale_y_continuous(expand = c(0,0), limits = c(0, 0.026), breaks = seq(0.012, 0.024, 0.004))

#indiaplots


#Estimate Carrying Capacity for India
india_ck <- pop_data_df %>% 
  filter(country == "India") %>% 
  lm(population ~ dNNdt, data =.)

india_ck
summary(india_ck)

brazil_ck <- pop_data_df %>% 
   filter(country == "Brazil") %>% 
  lm(population ~ dNNdt, data =.)

brazil_ck
summary(brazil_ck)
```
#### India
##### $Population = 2.252*10^9 - 7.226*10^{10}(\frac{dN}{Ndt})$
#####The carrying capacity (K) when $\frac{dN}{Ndt}$ = 0 is $2.252*10^9$ or 2,252,000,000 people.


#### Brazil
##### $Population = 2.607*10^8 - 6.392*10^9(\frac{dN}{Ndt})$
#####The carrying capacity (K) when $\frac{dN}{Ndt}$ = 0 is $2.607*10^8$ or 260,700,000 people. 


###C. Modify the plot to include year and food consumption
```{r, echo = FALSE, warning = FALSE}
country_plots_color <- ggplot(pop_data_mills_df, aes(x = pop_mills, y= dNNdt)) +
  geom_line()+
  facet_wrap(~country, scales = "free") + 
  labs(y = expression(paste("Per Capita Rate of Change ", (frac(dN, Ndt)))), x = "Population Size (Nt, Millions)", size = "Calories Consumed (Person/Day)") +
  theme_classic() +
  ggtitle("Per Capita Growth Rate versus Population Density of Five Countries") + geom_point(aes(color = year, size = calories_pd)) + scale_color_distiller(palette = "Purples", name = "Year", direction = 1) +
  theme(legend.position = "bottom",
        legend.box = "vertical")

country_plots_color


```


##Question 3

###A. Perform a multilinear regression of education and food consumption for all countries
```{r, echo = FALSE, include = FALSE}

# Make a corr plot
country_corr <- pop_data_df %>% 
  select(1:6)


country_corr <- cor(pop_data_df[2:6])
country_cp <- corrplot(country_corr,
                   type = "upper",
                   method = "ellipse",
                   tl.col = "darkorchid4",
                   tl.cex = 0.5)

# Create a linear model
pop_lm1 <- lm(dNNdt ~ years_ed + calories_pd, data = pop_data_df)
vif(pop_lm1)
AIC(pop_lm1)
summary(pop_lm1)
plot(pop_lm1)

# Create a linear model for all countries

#Brazil
brazil_lm <- pop_data_df %>% 
  filter(country == "Brazil")

brazil_lm1 <- lm(dNNdt ~ years_ed + calories_pd, data = brazil_lm)
vif(brazil_lm1)
AIC(brazil_lm1)
summary(brazil_lm1)
plot(brazil_lm1)

# India
india_lm <- pop_data_df %>% 
  filter(country == "India")

india_lm1 <- lm(dNNdt ~ years_ed + calories_pd, data = india_lm)
vif(india_lm1)
AIC(india_lm1)
summary(india_lm1)
plot(india_lm1)

# Japan
japan_lm <- pop_data_df %>% 
  filter(country == "Japan")

japan_lm1 <- lm(dNNdt ~ years_ed + calories_pd, data = japan_lm)
vif(japan_lm1)
AIC(japan_lm1)
summary(japan_lm1)
plot(japan_lm1)

# Mexico
mexico_lm <- pop_data_df %>% 
  filter(country == "Mexico")

mexico_lm1 <- lm(dNNdt ~ years_ed + calories_pd, data = mexico_lm)
vif(mexico_lm1)
AIC(mexico_lm1)
summary(mexico_lm1)
plot(mexico_lm1)

# South Korea
sk_lm <- pop_data_df %>% 
  filter(country == "South Korea")

sk_lm1 <- lm(dNNdt ~ years_ed + calories_pd, data = sk_lm)
vif(sk_lm1)
AIC(sk_lm1)
summary(sk_lm1)
plot(sk_lm1)



```

####All Countries:
Based on the linear regression
$\frac{dN}{Ndt}\ $ = $2.975*10^{-2}$ + $1.292*10^{-7}$(Calories (Person/Day)) - $2.95*10^{-2}$(Years Educated), education level is a significant predicator of per capita population growth rate (p<0.001).  However, food consumption, defined as the number of calories eaten per person, per day does not have a significant (p=0.549) effect on $\frac{dN}{Ndt}\ $.

All other factors remaining constant, for each unit increase in education level, $\frac{dN}{Ndt}\ $ significantly decreases by $2.95*10^{-2}$. 

All things other factors remaining constant, for each unit increase in calories per person, per day, $\frac{dN}{Ndt}\ $ increases by $1.292*10^{-7}$. 

###B. Perform a multilinear regression of education and food consumption for each country
```{r, echo = FALSE, results = 'asis'}
##### Create Tables

regression_table <- stargazer(pop_lm1, brazil_lm1, india_lm1, japan_lm1, mexico_lm1, sk_lm1,
                      title = "Linear Regression of Education and Food Consumption Results",
                      type = "html",
                      dep.var.labels = c("Per Capita Rate of Change (dN/Ndt)"),
                      column.labels = c("All", "Brazil", "India", "Japan", "Mexico", "South Korea"), covariate.labels = c("Years of Education", "Calories Consumed Per Person Per Day"), digits=6,
                      header = FALSE)


```

####Brazil:
Based on the linear regression
$\frac{dN}{Ndt}\ $ = $3.631*10^{-2}$ - $1.150*10^{-6}$(Calories (Person/Day)) - $2.420*10^{-3}$(Years Educated), education level is a significant predicator of per capita population growth rate (p<0.001).  Food consumption, though less impactful than education, also has a significant (p<0.05) effect on $\frac{dN}{Ndt}\ $.

All other factors remaining constant, for each unit increase in education level, $\frac{dN}{Ndt}\ $ decreases by $2.420*10^{-3}$. 

All things other factors remaining constant, for each unit increase in calories per person, per day, $\frac{dN}{Ndt}\ $ decreases by $1.150*10^{-6}$. 

####India
Based on the linear regression
$\frac{dN}{Ndt}\ $ = $2.600*10^{-2}$ - $4.433*10^{-7}$(Calories (Person/Day)) - $1.523*10^{-3}$(Years Educated), education level is a significant predicator of per capita population growth rate (p<0.001).  Food consumption, does not have a significant (p<0.05) effect on $\frac{dN}{Ndt}\ $ (p=0.793).

All other factors remaining constant, for each unit increase in education level, $\frac{dN}{Ndt}\ $ decreases by $1.523*10^{-3}$. 

All things other factors remaining constant, for each unit increase in calories per person, per day, $\frac{dN}{Ndt}\ $ decreases by $1.523*10^{-3}$. 

####Japan:
Based on the linear regression
$\frac{dN}{Ndt}\ $ = $5.061*10^{-2}$ - $1.978*10^{-6}$(Calories (Person/Day)) - $3.834*10^{-3}$(Years Educated), education level is a significant predicator of per capita population growth rate (p<0.001).  Food consumption, also has a significant effect on $\frac{dN}{Ndt}\ $ (p<0.05).

All other factors remaining constant, for each unit increase in education level, $\frac{dN}{Ndt}\ $ decreases by $3.834*10^{-3}$. 

All other factors remaining constant, for each unit increase in calories per person, per day, $\frac{dN}{Ndt}\ $ decreases by $1.978*10^{-6}$. 

####Mexico:
Based on the linear regression
$\frac{dN}{Ndt}\ $ = $5.888*10^{-2}$ - $5.306*10^{-6}$(Calories (Person/Day)) - $2.280*10^{-3}$(Years Educated), education level is a significant predicator of per capita population growth rate (p<0.001).  Calories eaten per person, per day simlilarly has a very significant effect on $\frac{dN}{Ndt}\ $ (p<0.001) in Mexico.

All other factors remaining constant, for each unit increase in education level, $\frac{dN}{Ndt}\ $ decreases by $2.280*10^{-3}$. 

All things other factors remaining constant, for each unit increase in calories per person, per day, $\frac{dN}{Ndt}\ $ decreases by $5.306*10^{-6}$. 

####South Korea:
Based on the linear regression
$\frac{dN}{Ndt}\ $ = $3.858*10^{-2}$ - $4.279*10^{-7}$(Calories (Person/Day)) - $2.757*10^{-3}$(Years Educated), education level is a significant predicator of per capita population growth rate (p<0.001).  Calories eaten per person, per day is not a statistically significant predictor of $\frac{dN}{Ndt}\ $ (p=0.0658).

All other factors remaining constant, for each unit increase in education level, $\frac{dN}{Ndt}\ $ decreases by $2.757*10^{-3}$. 

All other factors remaining constant, for each unit increase in calories per person, per day, $\frac{dN}{Ndt}\ $ decreases by $4.279*10^{-7}$. 



