---
title: "Assignment1_FINAL"
author: "Sandra Fogg"
date: "2/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, echo=FALSE}
# Load packages
library(tidyverse)
library(corrplot)
library(janitor)
library(lubridate)
library(ggimage)
library(kableExtra)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(extrafont)
library(ggrepel)
library(vcdExtra)
library(effsize)
library(leaflet)
library(car)
library(stargazer)
library(descr)
library(magrittr)
```

```{r, include=FALSE, echo=FALSE}
# Read csv data
pop_data_df <- read_csv("assig1_data.csv")
pop_data_df
pop_data_mills_df <- pop_data_df %>% 
  mutate(pop_mills = population/1000000)
pop_data_mills_df
```


###Question 1

####A. Describe how annual population data are used to calculate $\frac{dN}{Ndt}$. 
The expression $\frac{dN}{Ndt}$ represents the per capita rate change of a population (r). By taking the natural log of the annual population at time (t), dividing it by the number of individuals in the same population at time (t+1), and multiplying it by the expression $\frac{1}{(t+1)-t}$, the annual population data can be used to estimate per capita rate of change of the population.

####B. Provide the calculation used to determine $\frac{dN}{Ndt}$ for each country during the time period 1963 to 1964.

##### Brazil, $r= ln\left(\frac{81972001}{79602001})\right)$

##### India, $r= ln\left(\frac{486639001}{476632001})\right)$

##### Japan, $r= ln\left(\frac{96959001}{95929001})\right)$

##### Mexico, $r= ln\left(\frac{43052001}{41715001})\right)$

##### South Korea, $r= ln\left(\frac{27767001}{27138001})\right)$

```{r, include = FALSE, echo = FALSE}
brazil_r <- log(81972001/79602001)
brazil_r

india_r <- log(486639001/476632001)
india_r

japan_r <- log(96959001/95929001)
japan_r

mexico_r <- log(43052001/41715001)
mexico_r

southkorea_r <- log(27767001/27138001)
southkorea_r

set_colnames <- 

r_table <- pop_data_mills_df %>% 
  filter(year %in% c(1963,1964)) %>% 
  select(country, year, pop_mills) %>% 
  spread(year, pop_mills) %>% 
  set_colnames(value = c("Country", "N0", "Nt")) %>% 
  mutate(dNNdt_calc = log(Nt/N0),
         dNNdt = c(0.029338507, 0.020777870, 0.010679874, 0.031547910, 0.022913305))


r_table
```

```{r, echo= FALSE}
rate_table  <- r_table %>%
  select(Country, dNNdt_calc, dNNdt)
kable(rate_table, digits = 4, col.names = c("Country",  "Calculated dN/Ndt", "Actual dN/Ndt"),
      caption = "Calculated vs. Actual Per Capita Rate of Population Change by Country from 1963 to 1964") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "bordered")) 


```
####C. Calculate population rate change of one country from 1963 - 2004
```{r}

```

